\section{Measuring a Real World Example Processor}

The results contributed through this work relies on the existence of a method to
isolate and measure core voltage on a hardware implemented reference CPU. This
is made easy due to the $V_{core}$ separation on our development kit. In
\cite{rundehvatum2013exploring}, we conducted experiments to quantify the energy
cost of an instruction executing on a modern out-of-order mobile processor while
sampling at significantly lower rate. We were able to do this by completely
bypassing the memory hierarchy utilizing special hardware (\emph{fast-loop
mode}) and sampling a running average. Voltage drop over a shunt resistor such
as the one shown in \autoref{fig:shunt} was set in series with the ODROID-X2
development board is measured, and we calculate energy used in the processor
core. We isolate the energy used in the CPU (in which we are interested in) with
energy used for on-board peripherals by modifying the development board and
driving the CPU core with a separate power supply.

To derive the energy used in the processor, we first calculate the amount of
current flowing through the circuit. We assume that the shunt resistor voltage
drop is negligible compared the voltage drop over the ODROID, so that we have
1.3V on both sides of the resistor. Ohms law expresses current as

\begin{equation}
    I=\frac{U}{R}
    \label{eq:ohm}
\end{equation}

The voltage $U$ is being measured, while the restistance $R$ is a known fixed
value (12 m$\Omega$ in our experiments). Relating current to energy consumption
is done by applying the definition of electric power.

\begin{equation}
    P=U \cdot I
    \label{eq:power}
\end{equation}

Now, the current through the shunt resistor is the same as through the CPU core,
and the voltage is assumed to a fixed 1.3V.


