\pagestyle{empty}
\begin{abstract}

    \noindent Energy efficiency is currently one of the most challenging goals
    of modern computer design. High power density limits further performance
    growth, and energy efficiency affects both the power bill for
    supercomputers and battery lifetime for embedded devices. In this thesis we
    investigate energy consumption and architectural properties of an ARM
    Cortex-A9 and create a tool for estimating its power consumption through
    simulation.

    Instruction level energy consumption is determined through measurements and
    experiments done on real hardware, which are further mapped to architectural
    events found in the gem5 simulator. We write a tool which utilizes these
    events together with a simulator trace log and outputs a representation of
    energy consumption over time.

    Our method can be applied during the development process at the simulator
    level, while traditional methods typically involves hardware synthesis. The
    results shows that our tool can estimate energy consumption within 10~\% on
    general workloads, and is able to identify power consumption trends
    throughout a program.

\newpage
{\huge Old Abstract}\\
    \noindent Since the beginning of semiconductor technology, transistor size has
decreased and are still decreasing with a tremendous rate. Computer engineers
and architects have utilized higher transistor count to create more advance and
more dense computer chips, but are now facing problems with power consumption
and heat dissipation. In this thesis, an ARM-based SoC is studied in great
detail and is simulated using the gem5 architectural simulator. Using a genetic
algorithm, different power consuming events was mapped to a fixed energy
contribution.

Similar to measuring the execution of a program on a real processor using an
ammeter, we have built a tool called PET which provides the same for a
simulated program on unrealized hardware. PET parses trace logs from the
simulator and applies the energy contribution from each event. The output is
given as a textual or graphical timeline with current drain plotted at each data
point.

Comparing to other tools which does much the same, PET works much earlier in the
design stage, and can be applied without knowledge about process technology, RTL
or the physical layout of a chip. This tool can help engineers figure out how a new
feature or a change in an existing chip will influence energy efficiency.
Benchmarks and test data indicates that PET is able to predict power consumption
within 10~\% of the real value. Given the abstraction level at which PET works,
more accuracy would need further information. PET will give an indication of
better or worse, not an exact prediction of energy consumption.

\end{abstract}
