\chapter{Building a Power Estimation Tool}

The ultimate goal for this project is to model and estimate energy consumption
for not yet implemented computer architectures. This allows new ideas to be
prototyped and evaluated with respect to energy efficiency already at the design
stage, easing the process of building energy efficient hardware. These
evaluations can only serve as estimates and will doubtedly be truly accurate.
Nevertheless, it can be used to test specific workloads and applications on
specific processor configurations and evaluate ideas rapidly during the design
phase.

As we are already supplied with a computer architecture simulator capable of
tracing all sorts of hardware events, the next step is to extract power
information from these event logs. In this chapter we introduce PET, a Power
Estimation Tool. PET provides guided information about power usage for computer
architectures and represents a major piece of our problem solution. The chapters
to follow describe how it works along with its strengths and weaknesses.

\noindent PET is implemented in C++ using the Boost Library \cite{boostwebpage}. The source code for
PET and the rest of this project can be found at Github:

\begin{lstlisting}[language=bash,numbers=none]
git clone https://github.com/terjr/thesis.git
\end{lstlisting}

\input{pet/concept.tex}
\input{pet/approach.tex}
\input{pet/input.tex}
\input{pet/output.tex}
\input{pet/architecture.tex} % this ended up as a "gluing it all together"-section. OK?

