\begin{titlingpage}

\noindent
\begin{tabular}{@{}p{4cm}l}
    \textbf{Title:}     & \thetitle \\
    \textbf{Students:}  & Terje Runde \& Stian Hvatum \\
\end{tabular}

\vspace{4ex}
\noindent\textbf{Problem description:}
The SHMAC prototype is an ongoing research project within the Energy Efficient
Computing Systems (EECS) strategic research area. SHMAC is planned to run in an
FPGA and be an evaluation platform for research on heterogeneous multi-core
systems. Due to the Dark silicon effect, future computing systems are expected
to be power limited. The goal of the SHMAC project is to propose software and
hardware solutions for future power-limited heterogeneous systems.

The micro architecture level is an implementation of the Instruction Set
Architecture (ISA). Energy efficiency of an ISA is as such given by the chosen
micro architecture. To be able to take the "right" design choice to optimize for
energy efficiency, knowledge of energy and power for instruction types, e.g.,
instructions of type float, nop, copy, are needed.

The goal of this sub-project within the SHMAC platform is to gain knowledge of
energy/power consumption of different instruction types to be able to extract
information that can be used to improve the micro architecture design of
SHMAC-cores. This project will take a twofold approach; 1) Investigate the
power/energy consumption of simple benchmark programs on real hardware, i.e.
create benchmark programs and evaluate performance by measurements. 2)
Investigate the same benchmark programs in simulations as to ensure a good
understanding of the relation between measurements and simulated results.

\noindent The project will include:
\setitemize[0]{topsep=0pt,partopsep=0pt,parsep=0pt,itemsep=0pt}
\begin{itemize}
    \item Devising small benchmark programs, e.g., C or assembly, that isolate
    specific functions at the micro architecture level.  \item Run test on real
        hardware to collect data.
    \item Run tests in simulation to relate measurements to simulation results.
\end{itemize}

\noindent An ARM processor is going to be the target ISA for measurements and
simulations.
\vspace{6ex}

\noindent
\begin{tabular}{@{}p{4cm}l}
    \textbf{Supervisor:}            & \thesupervisor \\
    \textbf{Co-supervisor:}          & \thecosupervisor \\
\end{tabular}

\end{titlingpage}
