\section{CPU Modelling in gem5}

The revision of gem5 used for these experiments, changeset aaf017eaad7d,
contains a decent model for the ARM Cortex-A9 processor. Actually, the
implemented CPU is a generic out-of-order processor capable of running the ARMv7
ISA, which is equal to that of ARM Cortex-A9\cite{armtech}. The performance of
the ARM Cortex-A9 will differ between its implementations, and to make it even
harder, most vendors will not disclose details to the public. It has been
discussed among the gem5 users wherever the default configuration for this
processor actually is equal to any real processor\cite{a15maillist}, but the
general answer is that no one knows or are allowed to tell.

The goal of this project it to be able to estimate power for new architectures
that might or might not be close to existings architectures, but as a reference
for the implementation, a good model for a known processor is needed. The
hardware platform and the experimental setup is equalt to that in
\cite{rundehvatum2013exploring}, so gem5 must be adapted to simulate a Samsung
Exynos 4412P, or at least the CPU-part of it.  gem5 is made to be adapted, so
adaption is mostly done through editing python files. A new file was created in
\texttt{gem5/configs/common/Exynos\_4412P.py} and a few others where edited to
include this new processor definition. The processor definition used and other
modifications done to the original changeset aaf017eaad7d can be found in
\autoref{apx:gem5files}.

As a decent amount of implementation details often are keept secret from end users, it
is not trivial to model any commercial processor. To get as equal as possible performance
and event scheduling to the real hardware, parameters was fetched from around the internet,
and the simulator was then scripted to run with a wide variaty of configurations and
mutations of these, all for the ARM Out-of-order CPU model. When the run time performance was
as good as time permitted, the configuration was accepted and used further in the experiment.
This final configuration is the one printed in \autoref{gem5exynos4412p}. Sources suggesting
parameters for ARM Cortex-A9 or Samsung Exynos 4412P is listed in \autoref{fig:a9paramsources}

\begin{figure}
All web references was found between january and april 2014.
\begin{itemize}
    \item{\cite{blem2013detailed}}
    \item{\cite{butko2012accuracy}}
    \item{\cite{armtech}}
    \item{\url{http://en.wikipedia.org/wiki/Odroid}}
    \item{\url{http://en.wikipedia.org/wiki/Exynos}}
    \item{\url{http://www.geekland.co/ARM-Cortex-A9-Exynos-Quad-Core-4412-Android-Development-Board-GK-4412P.htm}}
    \item{\url{http://www.7-cpu.com/cpu/Cortex-A9.html}}
    \item{\url{http://www.arm.com/products/processors/cortex-a/cortex-a9.php?tab=Specifications}}
\end{itemize}
\caption{Sources suggesting parameters for ARM Cortex-A9 or Samsung Exynos 4412P}
\label{fig:a9paramsources}
\end{figure}
